---
title: "AAMC Revenue Report"
description: |
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(scales)
library(readxl)
library(DT)
library(rmarkdown)
```

## Revenue by Source 

Harvard has the highest revenue of all AAMC schools. It collects the most dollars in tuition, 
federal grants, and non-federal grants of all AAMC schools. It collects the third
most in gifts and endowment. 

```{r revenue_source, warning = FALSE, message=FALSE}
read_excel("aamc_revenue_source.xlsx") %>% 
  rename(school = `...1`) %>% 
  mutate(across(.cols = -school, ~str_replace_all(., "\\$|,", ""))) %>% 
  mutate(across(.cols = -school, as.numeric)) %>% 
  drop_na() %>% 
  rename(tuition = `Total Tuition and Fees Revenues`,
         federal_grants = `Total Federal Research Grants and Contracts`,
         other_grants = `Other Grants and Contracts`,
         support = `Total Government and Parent Support`,
         hospital = `Total Expenditures and Transfers from Hospital Funds`,
         gifts_endowment = `Total Revenues from Gifts and Endowment Funds`) %>% 
  arrange(desc(tuition)) %>% 
  rowwise() %>% 
  mutate(total = sum(across(.cols = -school))) ->a

a %>%
  group_by(1) %>% 
  summarize(across(-school, ~mean(.))) %>% 
  mutate(total_grants = federal_grants + other_grants) %>% 
  select(-`1`) %>% 
  mutate(school = "Average") %>% 
  select(school, tuition, federal_grants, other_grants, total_grants, support, hospital, gifts_endowment, total) %>% 
  mutate(across(-school, ~formatC(., format = "d", big.mark = ",")))-> a_average

datatable(a, rownames = FALSE)  %>% 
    formatCurrency(2:8, currency = "", interval = 3, mark = ",")

paged_table(a_average)
```

### Percent Revenue by Source

The proportion each revenue source contributes to the school's total revenue varies
widely among schools. Tution makes up 3% of Harvard's total revenue, while it makes
up 51% of Tufts'. The average percentage received from tuition is 13%.

Federal grantscontribute most to Harvard's total revenue, making up
47%. Of all AAMC schools, Harvard receives the second highest percentage, 89%, of its total
revenue from all grants, behind Mayo Clinic. The average percentage received from all grants is
roughly 30%. 

Harvard receives approximately 6% of its revenue from gifts and endowment, while 
Stanford receives approximately 19% from these, the second highest percentage after Oregon. 
The average percentage received from gifts and endowments is roughly 5%.
\
&nbsp;
\
&nbsp;


```{r}
a %>% 
  mutate(across(.cols = -school, ~./total)) %>% 
  mutate(total_grants = federal_grants + other_grants) %>% 
  select(school, tuition, federal_grants, other_grants, total_grants, support, hospital, gifts_endowment, total) -> a_percent

datatable(a_percent, rownames = FALSE) %>% 
  formatPercentage(2:8, digits = 2)

a_percent_avg <- a_percent %>% 
  group_by(total) %>% 
  summarize(across(-school, ~mean(.))) %>% 
  mutate(school = "Average") %>% 
  select(school, tuition, federal_grants, other_grants, total_grants, support, hospital, gifts_endowment, total) %>% 
  mutate(across(-school, ~str_c(as.character(round(.*100, digits =2)), "%")))

paged_table(a_percent_avg)
```

## Federal Research Grants and Contracts 

Harvard lags in recorded grants but leads in unrecorded grants. 

```{r grants, warning = FALSE, message=FALSE}
read_excel("aamc_grants.xlsx") %>% 
  rename(school = `...1`) %>% 
  mutate(across(.cols = -school, ~str_replace_all(., "\\$|,", ""))) %>% 
  mutate(across(.cols = -school, as.numeric)) %>% 
  drop_na() %>% 
  rename(direct_recorded = `Direct Federal Research Grants and Contracts - Recorded`,
         direct_unrecorded = `Direct Federal Research Grants and Contracts - Not Recorded`,
         indirect_recorded = `Indirect Federal Research Grants and Contracts - Recorded`,
         indirect_unrecorded = `Indirect Federal Research Grants and Contracts - Not Recorded`, 
         total = `Total Federal Research Grants and Contracts`) %>% 
  arrange(desc(direct_unrecorded))-> b

datatable(b, rownames = FALSE) %>% 
      formatCurrency(2:5, currency = "", interval = 3, mark = ",")
```

### Direct and Recorded by Percent

On average, schools have far more direct grants than indirect grants and far 
more recorded grants than unrecorded grants. Harvard follows the first trend. 
However, as noted above, Harvard has far more unrecorded grants than recorded 
grants. 

Note that percent recorded and percent unrecorded should add to 100, and likewise
for direct and indirect. 

```{r}
b_percent <- b %>% 
  mutate(recorded_percent = (direct_recorded + indirect_recorded) / total,
         unrecoreded_percent = (direct_unrecorded + indirect_unrecorded) /total, 
         direct_percent = (direct_recorded + direct_unrecorded) / total, 
         indirect_percent = (indirect_recorded + indirect_unrecorded) / total) %>% 
  select(school, contains("percent"))

 b_percent_avg <- b_percent %>% 
  group_by(1) %>% 
  summarize(across(-school, ~mean(., na.rm = TRUE))) %>% 
  select(-`1`) %>% 
   mutate(school = "Average") %>% 
   select(school, recorded_percent:indirect_percent) %>% 
   mutate(across(-school, ~str_c(as.character(round(.*100, digits =2)), "%")))
  
 
 datatable(b_percent, rownames = FALSE) %>% 
   formatPercentage(2:5, digits = 2)
  
 paged_table(b_percent_avg)
```

## Gifts and Endowments 

Harvard lags in gifts, 28th out of all medical schools, but is first in endowments. 

```{r gifts, warning = FALSE, message=FALSE}
read_excel("aamc_gifts.xlsx") %>% 
  rename(school = `...1`) %>% 
  mutate(across(.cols = -school, ~str_replace_all(., "\\$|,", ""))) %>% 
  mutate(across(.cols = -school, as.numeric)) %>% 
  drop_na() %>% 
  rename(gifts = `Total Revenues from Gifts`,
         endowment =`Total Revenues from Endowment Funds`, 
         total = `Total Revenues from Gifts and Endowment Funds`) %>% 
  arrange(desc(gifts)) -> c

datatable(c, rownames = FALSE) %>% 
  formatCurrency(2:4, currency = "", interval = 3, mark = ",")
```

### Percent Gifts and Endowments

Of the sum of gift and endowments, gifts make up approximately 14% while endowments
make up 86%, the tenth highest percentage of all AAMC schools. This is quite different from 
the mean. Johns Hopkins is one example of a school skewed in the other direction, 
with gifts making up 74% of its sum. 

```{r}
c_percent <- c %>% 
  mutate(percent_gifts = gifts / total, 
         percent_endowment = endowment / total) %>% 
  select(school, contains("percent")) 
  
c_percent_avg <- c_percent %>% 
  group_by(1) %>% 
  summarize(across(-school, ~mean(., na.rm = TRUE))) %>% 
  select(-`1`) %>% 
  mutate(school = "Average") %>% 
  select(school, percent_gifts, percent_endowment) %>% 
  mutate(across(-school, ~str_c(as.character(round(.*100, digits =2)), "%")))

 datatable(c_percent, rownames = FALSE) %>% 
   formatPercentage(2:3, digits = 2)
  
 paged_table(c_percent_avg)
```

## Government and Parent Support

Harvard exceeds the private school average in state and parent support. No money 
from any other form of support. 

```{r support, warning = FALSE, message=FALSE}
read_excel("aamc_support.xlsx") %>% 
  rename(school = `...1`) %>% 
  mutate(across(.cols = -school, ~str_replace_all(., "\\$|,", ""))) %>% 
  mutate(across(.cols = -school, as.numeric)) %>% 
  drop_na() %>% 
  rename(federal = `Total Federal Appropriations`,
         state_parent = `Total Adjusted State and Parent Support`,
         local = `Total Local Appropriations`,
         total = `Total Government and Parent Support`) %>% 
  filter(school %in% c("Private, All Regions Average", "Harvard")) -> d

datatable(d, rownames = FALSE) %>% 
  formatCurrency(2:5, currency = "", interval = 3, mark = ",")
```

## Hospital and Medical School Programs - by Hospital Type

Harvard receives only from the "Total Expenditures and transfers from Other Affiliated Hospital Funds" category and lags behind the private school average in all.

```{r hospital_type, warning = FALSE, message=FALSE}
read_excel("aamc_hospital_type.xlsx") %>% 
  rename(school = `...1`) %>% 
  mutate(across(.cols = -school, ~str_replace_all(., "\\$|,", ""))) %>% 
  mutate(across(.cols = -school, as.numeric)) %>% 
  drop_na() %>% 
  rename(university_hospital = `Total Expenditures and transfers from University Hospital Funds`,
         va_hospital = `Total Expenditures and transfers from VA Hospital Funds`,
         other_hospital = `Total Expenditures and transfers from Other Affiliated Hospital Funds`,
         total = `Total Expenditures and Transfers from Hospital Funds`) %>% 
  filter(school %in% c("Private, All Regions Average", "Harvard")) -> e

datatable(e, rownames = FALSE) %>% 
  formatCurrency(2:5, currency = "", interval = 3, mark = ",")
```

## Hospital and Medical School Programs - by Recipient of Funds

Harvard only funds strategic support for Medical School Programs. The sum 
is lower than the private school average.

```{r hospital_programs, warning = FALSE, message=FALSE}
read_excel("aamc_hospital_programs.xlsx") %>% 
  rename(school = `...1`) %>% 
  mutate(across(.cols = -school, ~str_replace_all(., "\\$|,", ""))) %>% 
  mutate(across(.cols = -school, as.numeric)) %>% 
  drop_na() %>% 
  rename(services = `Total Purchased Services`,
         stipends = `Total Housestaff Stipends`,
         clinical_investments = `Total Hospital Investments in the Clinical Enterprise`,
         strategic_support = `Total Strategic Support for Medical School Programs`,
         total = `Total Expenditures and Transfers from Hospital Funds`) %>% 
  filter(school %in% c("Private, All Regions Average", "Harvard")) -> f 

datatable(f, rownames = FALSE) %>% 
  formatCurrency(2:5, currency = "", interval = 3, mark = ",")
```

## Market Value of Endowment by Type

Harvard, along with many other schools, lacks breakdown for endowment. It has the highest total, 
in accordance with what we have seen above.

```{r aamc_endowment, warning = FALSE, message=FALSE}
# I don't understand the total column, although it appears to be that way in 
# the original. 

g <- read_excel("aamc_endowment_value.xlsx") %>% 
  rename(school = `...1`) %>% 
  mutate(across(.cols = -school, ~str_replace_all(., "\\$|,", ""))) %>% 
  mutate(across(.cols = -school, as.numeric)) %>% 
  drop_na(school) %>% 
  rename(quasi_endowment = `Market Value of Quasi-endowments`,
         term_endowment = `Market Value of Term Endowments`,
         pure_endowment = `Market Value of True/pure Endowments`,
         other_endowment = `Market Value of Other Endowments`, 
         total = `Total Endowment (only if breakdowns are unavailable)`) %>% 
  rowwise() %>% 
  mutate(total = case_when(is.na(total) == TRUE ~  sum(quasi_endowment, term_endowment, pure_endowment, other_endowment, na.rm = TRUE),
    total == 0 ~ sum(quasi_endowment, term_endowment, pure_endowment, other_endowment, na.rm = TRUE),
    total > 0 ~ total)) 

datatable(g, rownames = FALSE) %>% 
  formatCurrency(2:6, currency = "", interval = 3, mark = ",")
```

## Federal Negotiated Facilities & Administration Rate

Harvard has the highest F&A rate, together with Cornell-Weill and NYU-Grossman.

```{r fa, warning = FALSE, message=FALSE}
read_excel("aamc_fa.xlsx") %>% 
  rename(school = `...1`) %>% 
  mutate(across(.cols = -school, as.numeric)) %>% 
  drop_na(school) %>% 
  rename(fa_rate = `Federal Negotiated F&A rate for on campus organized research`) %>% 
  arrange(desc(fa_rate)) -> h

datatable(h, rownames = FALSE) %>% 
  formatPercentage(2, digits = 1)
```

## Distribution of the Facilities & Administration Costs

Harvard has the third highest proportion of costs in the "other" category. It is exceeded only
by Oakland Beaumont and Zucker Hofstra Northwell. 

```{r fa_distribution, warning = FALSE, message=FALSE}
read_excel("aamc_fa_distribution.xlsx") %>% 
  rename(school = `...1`) %>% 
  mutate(across(.cols = -school, ~str_replace_all(., "%", ""))) %>% 
  mutate(across(.cols = -school, as.numeric)) %>% 
  mutate(across(.cols = -school, ~ . / 100)) %>% 
  drop_na(school) %>% 
  rename(medical_school = `Percent retained by medical school`,
        health_admin= `Percent retained by health sciences administration`,
        university_admin = `Percent retained by central university administration`, 
        state = `Percent retained by state`,
        other = `Percent retained by other`) %>% 
  arrange(desc(other)) -> i

datatable(i, rownames = FALSE) %>% 
  formatPercentage(2:6, digits = 1)
```

## Number of Faculty

Harvard has the most clinical faculty. Harvard has the 7th largest preclincial 
faculty.

```{r faculty, warning = FALSE, message= FALSE}
read_delim("aamc_faculty.csv", delim = ";") %>% 
  rename(school = X1) %>% 
  select(-c(X5:X8)) %>% 
  mutate(across(.cols = -school, ~str_replace_all(., ",", ""))) %>% 
  mutate(across(.cols = -school, as.numeric)) %>% 
  drop_na(school) %>% 
  rename(pre_clinical = `Total Full-Time Basic Sciences Faculty`,
         clinical = `Total Full-Time Clinical Faculty`,
         total = `Total Full-Time Faculty`)  -> j 
datatable(j, rownames = FALSE)
  
  
```